Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2012-03-08 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2012:03:08	12:34:44	(Thu 8.Mar)

\begin{itemize}
\item Sell it!
\item As briefly as possible, introduce beamforming.
\item Adaptive beamformers
\item Challenges in active sonar systems. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive.
\item LCA is like DAS, only we test several windows and choose the one that yield the lowest beamformer output power. Ultrasound \cite{Synnevag2008}, active sonar \cite{Buskenes2011}\cite{Blomberg2011}\cite{Blomberg2012}\todo{who was first?}
\item Outline: Start with Capon's response when means for robustification are added.
\end{itemize}

@@

\begin{itemize}
\item Why beamforming?
\item Adaptive beamformers
\item Challenges in active sonar systems. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive.
\item LCA is like DAS, only we test several windows and choose the one that yield the lowest beamformer output power. Ultrasound \cite{Synnevag2008}, active sonar \cite{Buskenes2011}\cite{Blomberg2011}\cite{Blomberg2012}\todo{who was first?}
\item Outline: Start with Capon's response when means for robustification are added.
\end{itemize}

Outline:
\begin{itemize}
\item Capon's resonse when applying robustification
\item Choice of window functions makes little difference.
\item Steering and mainlobewidths have outer bounds.
\item Beamspace?
\item Chosen window plots - what may they tell us? Variance intensity values when using various windows.
\item Assymmetric windows?
\end{itemize}

\begin{align}
z[n] = \sumb{m=0}{M-1} w_m[n]^*x_m[n-\Delta_m] = \w\H[n]\x[n-\Delta_m]
\end{align}

\section{Methods}

Basically, we are working with a practical implementation of the Capon beamformer that computes a set of weights $\vec w$ for every single sample $n$ by solving:
\begin{gather*}
\vec w[n] = \frac{\hat{\mat R}\,\!^{-1}[n]\vec a}{\vec a\H\hat{\mat R}\,\!^{-1}[n]\vec a} = \frac{\text{\raisebox{1.9pt}{$\vec\chi$}}[n]}{\vec a\H\text{\raisebox{1.9pt}{$\vec\chi$}}[n]}
\end{gather*}%
where
% \newcommand\X{\text{\raisebox{2pt}{$\vec\chi$}}}
\begin{gather*}
\text{\raisebox{1.9pt}{$\vec\chi$}} = \hat{\mat R}\,\!^{-1}\vec a \qquad\qquad\text{is the solution to}\qquad\qquad \hat{\mat R}\text{\raisebox{1.9pt}{$\vec\chi$}} = \vec a.
\end{gather*}
Here $\vec a$ is a \emph{steering vector} that if set to $\vec 1$ steers towards broadside, and $\hat{\mat R}$ is an estimate of the spatial covariance matrix, often computed as
\begin{gather*}
r_{ij}[n] = \frac{1}{K(2N_{\text{avg}}+1)}\sumb{k=0}{K-1} \sumb{n'=n-N_{\text{avg}}}{n+N_{\text{avg}}} x_{i+k}[n']\,x_{j+k}[n'],
\end{gather*}
where $r_{ij}$ is the $(i,j)$'th element of the covariance matrix, $K$ is the number of subarrays, $2N_{\text{avg}}+1$ is the number of samples to perform time averaging over, and $\vec x_i[n']$ is the data recorded by the $i$'th sensor at sample $n'$. \emph{Diagonal loading} is usually also applied to constrain the adaptability and ensure numerical stability, but this operation is insignificant in the context of performance and will be ignored.

, and maximum suppression of while ensuring that the beamformer digitally  before each of the pixels are estimated one at a time. The resolution and contrast of such a system will depend on the systems spatial response, which ideally should be narrow  be very sharp in the desired direction its ability to achieve  fundamental principle of forming a sonar image is to record the received wavefield,

image quality of a phased array sonar imaging systems depend on  the choice of weights to apply to each of the sensors are crucial.

A modern phased array imaging system may be thought of as a spatial filter. To achieve the best possible performance, the

resolution and contrast

Adaptive beamformers have only recently been introduced in active sonar imaging. For a while they were considered unsuited for this purpose because the backscattered signal is largely correlated with the

An attactive trait of the LCA is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.

An attactive trait of the LCA\todo{break at something} is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.

An attactive trait of the LCA is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.

\hfill mds

\hfill January 11, 2007

\subsection{Subsection Heading Here}
Subsection text here.

@@

\appendices
\section{Proof of the First Zonklar Equation}
Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgement
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

>>>>

\begin{itemize}
\item Sell it!
\item As briefly as possible, introduce beamforming.
\item Adaptive beamformer potential
\item Why active sonar make it hard. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive.
\item LCA is like DAS, only we test several windows and choose the one that yield the lowest beamformer output power. Ultrasound \cite{Synnevag2008}, active sonar \cite{Buskenes2011}\cite{Blomberg2011}\cite{Blomberg2012}\todo{who was that guy in the 80ties?}
\item Outline: Start with Capon's response when means for robustification are added.
\end{itemize}

@@

% \begin{itemize}
% \item Why beamforming?
% \item Adaptive beamformers
% \item Challenges in active sonar systems. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive.
% \item LCA is like DAS, only we test several windows and choose the one that yield the lowest beamformer output power. Ultrasound \cite{Synnevag2008}, active sonar \cite{Buskenes2011}\cite{Blomberg2011}\cite{Blomberg2012}\todo{who was first?}
% \item Outline: Start with Capon's response when means for robustification are added.
% \end{itemize}
% 
% 
% Outline:
% \begin{itemize}
% \item Capon's resonse when applying robustification
% \item Choice of window functions makes little difference.
% \item Steering and mainlobewidths have outer bounds.
% \item Beamspace?
% \item Chosen window plots - what may they tell us? Variance intensity values when using various windows.
% \item Assymmetric windows?
% \end{itemize}
% 
% 
% \begin{align}
% z[n] = \sumb{m=0}{M-1} w_m[n]^*x_m[n-\Delta_m] = \w\H[n]\x[n-\Delta_m]
% \end{align}
% 
% 
% \section{Methods}
% 
% Basically, we are working with a practical implementation of the Capon beamformer that computes a set of weights $\vec w$ for every single sample $n$ by solving:
% \begin{gather*}
% \vec w[n] = \frac{\hat{\mat R}\,\!^{-1}[n]\vec a}{\vec a\H\hat{\mat R}\,\!^{-1}[n]\vec a} = \frac{\text{\raisebox{1.9pt}{$\vec\chi$}}[n]}{\vec a\H\text{\raisebox{1.9pt}{$\vec\chi$}}[n]}
% \end{gather*}%
% where
% % \newcommand\X{\text{\raisebox{2pt}{$\vec\chi$}}}
% \begin{gather*}
% \text{\raisebox{1.9pt}{$\vec\chi$}} = \hat{\mat R}\,\!^{-1}\vec a \qquad\qquad\text{is the solution to}\qquad\qquad \hat{\mat R}\text{\raisebox{1.9pt}{$\vec\chi$}} = \vec a.
% \end{gather*}
% Here $\vec a$ is a \emph{steering vector} that if set to $\vec 1$ steers towards broadside, and $\hat{\mat R}$ is an estimate of the spatial covariance matrix, often computed as
% \begin{gather*}
% r_{ij}[n] = \frac{1}{K(2N_{\text{avg}}+1)}\sumb{k=0}{K-1} \sumb{n'=n-N_{\text{avg}}}{n+N_{\text{avg}}} x_{i+k}[n']\,x_{j+k}[n'],
% \end{gather*}
% where $r_{ij}$ is the $(i,j)$'th element of the covariance matrix, $K$ is the number of subarrays, $2N_{\text{avg}}+1$ is the number of samples to perform time averaging over, and $\vec x_i[n']$ is the data recorded by the $i$'th sensor at sample $n'$. \emph{Diagonal loading} is usually also applied to constrain the adaptability and ensure numerical stability, but this operation is insignificant in the context of performance and will be ignored.
% 
% , and maximum suppression of while ensuring that the beamformer digitally  before each of the pixels are estimated one at a time. The resolution and contrast of such a system will depend on the systems spatial response, which ideally should be narrow  be very sharp in the desired direction its ability to achieve  fundamental principle of forming a sonar image is to record the received wavefield, 
% 
% image quality of a phased array sonar imaging systems depend on  the choice of weights to apply to each of the sensors are crucial. 
% 
% A modern phased array imaging system may be thought of as a spatial filter. To achieve the best possible performance, the 
% 
% resolution and contrast 
% 
% Adaptive beamformers have only recently been introduced in active sonar imaging. For a while they were considered unsuited for this purpose because the backscattered signal is largely correlated with the 
% 
% 
% 
% We have implemented a Low Complexity Adaptive (LCA) beamformer, which adaptively selects a window from a predefined set. The set is comprised of windows that are typical solutions found by the Minimum Variance method. The LCA beamformer was tested using simulated and experimental data from the Kongsberg Maritime HISAS 1030 sonar. On a simulated scene with speckle, highlight and shadow, the beamformer offered better lateral edge definition compared to the MV beamformer, and speckle intensity and shape comparable to DAS and MV beamformers. These results were verified by the experimental data.
% 
% An attactive trait of the LCA is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.
% 
% We have implemented a Low Complexity Adaptive (LCA) beamformer, which adaptively selects a window from a predefined set. The set is comprised of windows that are typical solutions found by the Minimum Variance method. The LCA beamformer was tested using simulated and experimental data from the Kongsberg Maritime HISAS 1030 sonar. On a simulated scene with speckle, highlight and shadow, the beamformer offered better lateral edge definition compared to the MV beamformer, and speckle intensity and shape comparable to DAS and MV beamformers. These results were verified by the experimental data.
% 
% An attactive trait of the LCA\todo{break at something} is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.
% 
% We have implemented a Low Complexity Adaptive (LCA) beamformer, which adaptively selects a window from a predefined set. The set is comprised of windows that are typical solutions found by the Minimum Variance method. The LCA beamformer was tested using simulated and experimental data from the Kongsberg Maritime HISAS 1030 sonar. On a simulated scene with speckle, highlight and shadow, the beamformer offered better lateral edge definition compared to the MV beamformer, and speckle intensity and shape comparable to DAS and MV beamformers. These results were verified by the experimental data.
% 
% An attactive trait of the LCA is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.
% 
% \hfill mds
%  
% \hfill January 11, 2007
% 
% \subsection{Subsection Heading Here}
% Subsection text here.

@@

%%%%%%%%%%%%%%%%%%                              ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% DOCUMENT APPENDICES %
%%%%%%%%%%%%%%%%%%                              ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\appendices

% use section* for acknowledgement
\ifCLASSOPTIONcompsoc
  \section*{Acknowledgments}
\else
  \section*{Acknowledgment}
\fi
